Malacology: A Programmable Storage System Built on Ceph

### 2016-05-09 paper/paper.tex
sec 2 updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
Adds Section 4

\usepackage{times}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
%\usepackage{biblatex}
%\bibliography{paper}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Malacology: A Programmable Storage System Built on Ceph},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

@@

The object interface facility provided by Ceph is more than an
proof-of-concept active storage implementation.  Figure
\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users. Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

@@

@@

To enable the programmability of Ceph subsystems, Malacology relies
heavily on interface classes for introducing new functionality into the
MDSs and OSDs. Malacology introduces 3 components to manage
object/metadata interfaces:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  additional interface hooks and classes for OSDs and MDSs using the
  existing Ceph \texttt{cls} infrastructure
\item
  durable interface classes using the Ceph object store
\item
  versioned and consistent interfaces using Ceph's consensus service
\end{enumerate}

In the following section, we will discuss the existing Ceph
infrastructure and the changes necessary for Malacology. Our focus will
be to demonstrate the viability and usefulness of the Malacology
appraoch with example implementations that due to limited time for the
submission of this paper are not necessarily fully optimized.

\subsection{Additional Interface Hooks and
Classes}\label{additional-interface-hooks-and-classes}

Malacology adds new hooks to Ceph and a new interface class for both the
MDS and OSD that uses Lua~\cite{ierusalimschy:lua15,ierusalimschy:spe96}. Our
framework has added a mechanism for defining and running object and
metadata balancer classes using Lua. Our Lua bindings expose functions
and symbols both ways; the host program can call functions defined in
Lua and the Lua scripts can call functions defined in native C++. These
bindings are merged upstream. We choose Lua for 4 reasons: performance,
portabability, size, and and security.

Lua is a fast scripting language. It was designed to be an embedded
language and the LuaJIT virtual machines boasts near-native performance~\cite{pall:luajit15,grawinkel_pdsw12}. Lua is frequently
used in game engines to set policies but we use it here because most of
the user-defined object classes in Ceph are policies as well (see
\S\ref{table:objclasses}). Separating policy from mechanism is a driving
factor in using Lua.

\begin{figure}[htbp]
\centering
\includegraphics{figures/programmability-framework.png}
\caption{Clients add functionality by injecting Lua scripts with a new
monitor (M) command. After the ``proposal'' (using PAXOS terminology) is
accepted by the cluster of monitors, the script is redundantly and
consistently distributed to OSDs. \label{fig:programmability-framework}}
\end{figure}

@@

Our framework has a mechanism for defining and running object and
metadata balancer classes with scripting. Scripting is useful in large
systems for 3 reasons: - accelerates deployment: tweaking policies, like
changing a threshold or metric calculation, does not require a full
recompile or system start-up. - transparency: if the bindings carefully
expose the correct metrics and functions, the user does not need to
learn the internal functions, variables, or types of the system. -
portability: scripts can be exchanged across platforms and system
versions without compatibility issues.

This effectively decouples policy from mechanism, which helps future
designers explore trade-offs and isolates policy development from
code-hardened systems. Decoupling policy mechanism is not a new
technique but designing the storage system with programmability as a
first-class citizen is credits policy/mechanism separation as the
biggest reason for the success of the block allocation mechanisms in the
Fast File System~\cite{mckusick:fast2015-FFS} yet there
are no mechanisms for transparently exposing and modifying the
underlying logic in modern file systems.

Malacology has Lua bindings that expose functions and symbols both ways;
the host program can call functions defined in Lua and the Lua scripts
can call functions defined in native C++. These bindings are merged
upstream. We choose Lua for 4 reasons: performance, portabability, size,
and security.

\subsubsection{Using Lua}\label{using-lua}

Lua is a performs well as an embeddable scripting language. Although it
is an interpreted language, the bindings are designed to make the
exchange of variables and functions amongst languages straightforward.
The LuaJIT virtual machine boasts near-native performance, making it a
logical choice for scriptability frameworks in systems research~\cite{neto:dls14-luaos}. In storage systems, it has been effectively
used both on \cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
\cite{sevilla:sc15-mantle} the critical path, where performance is
important. Lua is frequently used in game engines to set policies but we
use it here because most of the user-defined classes in Ceph are
policies as well! We do not want to provide specific implementations,
like pulling data from objects or transferring them over the network,
but instead strive to say \emph{what to do} with the data once we have
it. Separating policy from mechanism is a driving factor in using Lua.

@@

The other two advantages, that Lua leaves a small memory footprint and
that is is secure, are discussed at length in~\cite{ierusalimschy_programming_2006,neto:dls14-luaos}. In short, Lua
provides relatively robust sandbox capabilities for protecting the
system against some malicious cases, but does not protect the system
from poor policies. Since Malacology does not use these features (yet),
we will avoid discussion here.

\subsubsection{Generalizing the Lua VM}\label{generalizing-the-lua-vm}

@@

Both the object and balancer interfaces define functions and attach them
to the core Lua sandbox. For example, the Lua object storage interface
class attaches data, extended attribute, object map, and object version
functions, to the the Lua core functions; the exact functions are shown
in Listing\textasciitilde{}\ref{lst:attach_functions}. The advantages of
this is that we avoid duplicating code, we provide a framework for
putting Lua code in other parts of the system, and we remove components
and APIs that are too integrated with the OSD.

\subsubsection{Generalizing the Class
Handler}\label{generalizing-the-class-handler}

Our framework can also load Lua interfaces from the local file system --
the same technique as the C/C++ object interfaces. First, the OSD looks
for C/C++ shared library. If the OSD cannot find the file, it looks for
a Lua script of the same name. The script is read into a string of the
C++ object class; this string is later forwarded to the native Lua
class. When a requests comes in, the method name is passed with the
\texttt{exec()} function and the corresponding function in the Lua class
is called. Now clients need to only send their Lua object class once and
the OSDs will store them locally. Also, clients can execute any Lua
handler they want.

@@

For balancer interfaces, clients put balancers into RADOS and the CephFS
metadata balancer invokes the operations in the user-defined class
remotely on the MDS.

{[}7:31{]} Nothing would stop C++ from being stashed in objects and
recompiled on whatever platform they are being loaded on dynamically
(like a DB compiles each query). There just hasn't been a need for such
a feature. The next blog post, which I didn't at all allude to in the
one you are reading, shows how the Lua scripts can be put into the OSD
map and then monitors manage and version the script, distributing them
to each OSD. Stored in objects vs stored in the OSD map is a debate, I
guess. I am not sure which makes more sense.

@@

We added a command,
\texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}},
that the user uses to inject the Lua object interface into the cluster.
By leveraging the monitor daemons, we get the consistency from the
monitors' version management, the distribution from the monitors' data
structures (which are already distributed), and the durability from the
robustness of the monitors and persistence with Paxos. The
implementation uses the placement group pool data structure which
maintains the policies (\emph{e.g.}, erasure coding) and organization
details (\emph{e.g.}, snapshots) for each pool. While stuffing the
information into the monitor map, the structure that describes the the
monitor topology, was an option, the placement group pool allows finer
grainer control over different typoes of data. In regards to the
consistency, each time the user injects a new Lua object class it enters
the monitor quorum as a proposal; updates to the placement group pool
need to wait until accepted, at which point the state will be propogated
and versioned. We hooked up the monitor command to the \texttt{cls\_lua}
class by having the OSD pull the script from the placement group instead
of trying to dynamically load the shared library with \texttt{dlopen()}
or the Lua script.

Advantages:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  lets us store/manage interfaces
\item
  does most of the work (versioning, consistency, durabiltiy) for us
\item
  gives a new abstraction for dealing/mutating interfaces
\end{itemize}

@@

\subsection{Specifying the Lua Class}\label{specifying-the-lua-class}

Maintain versions and consistency

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  cls-lua branch: write object classes in Lua
\item
  lua-rados: talk to RADOS with Lua
\item
  cls-client: use lua-rados + cls-lua branch to send Lua object classes
  to OSDs equipped with LuaJIT VM
\end{itemize}

>>>>

\usepackage{fullpage}
\usepackage{times}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{graphicx}
\usepackage[small]{caption}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Malacology: A Programmable Storage System Built on Ceph},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

@@

The object interface facility provided by Ceph is more than an
a proof-of-concept active storage implementation.  Figure
\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users.
Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/implementation.png}
\caption{Malacology re-uses Ceph subsystems to manage interface classes.
command. After the proposal is accepted by the cluster of monitors, the
script is redundantly and consistently distributed to OSDs.
\label{fig:implementation}}
\end{figure}

@@

To enable the re-use of Ceph subsystems, Malacology relies heavily on
interface classes for introducing new functionality into the MDSs and
OSDs. Malacology introduces 3 components to manage object/metadata
interfaces:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  additional interface hooks and classes for OSDs and MDSs using the
  existing Ceph \texttt{cls} infrastructure
\item
  durable interface classes using the Ceph object store
\item
  versioned and consistent interfaces using the Ceph MONs
\end{enumerate}

Figure \ref{implementation} shows how interfaces can be loaded into
either the MDS or OSD (1), how interfaces are persisted to object
storage (2), and how they are versioned by the MONs (3). In the
following section, we will discuss the existing Ceph infrastructure and
the changes necessary for Malacology. We emphasize here that these
implementations are not necessarily optimized but they help demonstrate
the power of programmability.

\subsection{Additional Interface
Frameworks}\label{additional-interface-frameworks}

@@

Although these OSD interfaces are powerful, the current implementation
has drawbacks. OSD interfaces are written in C/C++ and compiled into a
shared library, so the developer must account for different target
architectures. Second, C/C++ provides more functionality than is
necessary since the snippets of interface code mainly set policies and
perform simple operations. Finally, developing these interfaces in C/C++
has high overhead. The developer must learn how the OSD dynamically
loads the shared libraries (\emph{e.g.}, OSDs rely on a strict naming
convention to find shared libraries), how to get their interfaces
compiled using the Ceph \texttt{make} system, and how to debug issues
that are not related to their specific interface (\emph{e.g.}, the OSD
cannot find the shared library) - this learning curve is unacceptable
for non-Ceph developers, especially since most interfaces are one-off
solutions specific to their applications.

\subsubsection{New Interface Classes}\label{new-interface-classes}

Our framework has added a mechanism for defining and running object and
metadata balancer classes using the Lua scripting language. Scripting is
useful in large systems for 3 reasons:

\begin{itemize}
\item
  accelerates deployment: tweaking policies, like changing a threshold
  or metric calculation, does not require a full recompile or system
  start-up.
\item
  transparency: if the bindings carefully expose the correct metrics and
  functions, the user does not need to learn the internal functions,
  variables, or types of the system.
\item
  portability: scripts can be exchanged across platforms and system
  versions without compatibility issues.
\end{itemize}

Our Lua bindings expose functions and symbols both ways; the host
program can call functions defined in Lua and the Lua scripts can call
functions defined in native C++. These bindings are merged upstream. Lua
is frequently used in game engines to set policies but we use it here
because most of the user-defined classes in Ceph are policies as well!
We do not want to provide specific implementations, like pulling data
from objects or transferring them over the network, but instead strive
to say \emph{what to do} with the data once we have it. Separating
policy from mechanism is a driving factor in using Lua. We choose Lua
for 4 reasons: performance, portabability, size, and security.

Lua performs well as an embeddable scripting language. Although it is an
interpreted language, the bindings are designed to make the exchange of
variables and functions amongst languages straightforward. The LuaJIT
virtual machine boasts near-native performance, making it a logical
choice for scriptability frameworks in systems research
\autocite{neto:dls14-luaos}. In storage systems, it has been effectively
used both on \autocites[
]{grawinkel:pdsw2012-lua}{watkins2013:bdmc13-in-vivo} and off
\autocite{sevilla:sc15-mantle} the critical path, where performance is
important.

@@

The other two advantages, that Lua leaves a small memory footprint and
that it is secure, are discussed at length in \autocites[
]{ierusalimschy_programming_2006}{neto:dls14-luaos}. In short, Lua
provides relatively robust sandbox capabilities for protecting the
system against some malicious cases, but does not protect the system
from poor policies. Since Malacology does not use these features (yet),
we will avoid discussion here.

\subsection{New Interface Hooks}\label{new-interface-hooks}

Malacogy adds interface hooks to the MDS in the load balancing and
capability subsystems. In contrast to the Mantle implementation, which
had a large function hard-coded into the balancer subsystem, Malacology
embeds code in the MDS using OSD subsystems. Ceph has a whole
infrastructure for getting dynamic code into the OSD. The
\texttt{ClassHandler} class safely opens object interfaces, executes
commands defined by the shared library, and fails gracefully with
helpful errors if anything goes wrong. Instead of injecting strings
directly into Ceph (like we did with the original Mantle
implementation), the \texttt{ClassHandler} class safely opens shared
code, even code with dependencies that are not in Ceph itself (e.g.,
Lua).

With the \texttt{ClassHandler}, the MDS gets the safety and robustness
of loading dynamic code, the ease of transferring state between object
interfaces and the OSD internals, integration with testing and
correctness suites, and \texttt{struct}s for interface data and
handlers. Using the OSD class handler hooks also gives the MDS the
ability to load interfaces from the local file system.

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/cls-osd-mds.png}
\caption{To generalize the LuaVM for use in both the OSD and MDS,
Malacology detaches the core functionality into a Lua core module. The
VM is dynamically loaded first (1) and then interfaces can execute (2).
\label{fig:cls-osd-mds}}
\end{figure}

Both the object and balancer interfaces define functions and attach them
to the core Lua sandbox. For example, the Lua object storage interface
class attaches data, extended attribute, object map, and object version
functions, to the Lua core functions, as shown in Figure
\ref{fig:cls-osd-mds}. The advantages of this is that we avoid
duplicating code, we provide a framework for putting Lua code in other
parts of the system, and we remove components and APIs that are too
integrated with the OSD.

@@

Traditionally, object storage interfaces are compiled into shared
libraries and loaded into a running OSD daemon using \texttt{dlopen()}.
Because the shared library has to link against symbols found in the
running executable, the interfaces are stored on the local file systems
of each OSD so they can be versioned and distributed alongside the same
Ceph binaries. This policy is for safety, since the OSDs could be on
different distros that provide different versioning capabilities. Also,
this approach allows bug fixes to co-evolve with the rest of the Ceph
installation.

Because of the portability of Lua and because the Lua virtual machine is
compiled and loaded as a shared library, the MDS/OSD can store the
balancers in a more redundant location. For our implementation, we store
balancers in RADOS and the daemons first search RADOS, then the local
file system for their interface classes.

@@

We added two new commands to Ceph command line client so the user can
inject interface versions into the cluster::

\noindent \texttt{ceph\ osd\ pool\ set-class\ \textless{}pool\textgreater{}\ \textless{}class\textgreater{}\ \textless{}script\textgreater{}}

\noindent \texttt{ceph\ mds\ set\ lua\_balancer\_class\ \textless{}version\textgreater{}}

By leveraging the monitor daemons, our services get the consistency from
the monitors version management, the distribution from the monitors data
structures (which are already distributed), and the durability from the
robustness of the monitors and persistence with Paxos. The
implementation uses the placement group pool data structure which
maintains the policies (\emph{e.g.}, erasure coding) and organization
details (\emph{e.g.}, snapshots) for each pool. While stuffing the
information into the monitor map, the structure that describes the the
monitor topology, was an option, the placement group pool allows finer
grain control over different types of data.

In regards to the consistency, each time the user injects a new Lua
object class it enters the monitor quorum as a proposal; updates to the
placement group pool need to wait until accepted, at which point the
state will be propogated and versioned. We hooked up the monitor command
to the \texttt{cls\_lua} class by having the OSD pull the script from
the placement group instead of trying to dynamically load the shared
library with \texttt{dlopen()} or the Lua script.

@@
